---
title: "Exercises 1"
author: "Abby Johnson"
date: "2/11/2022"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(rsample) 
library(caret)
library(modelr)
library(parallel)
library(foreach)

ABIA <- read.csv("~/Desktop/UT/Statistical Learning/ABIA.csv")

sclass <- read.csv("~/Desktop/UT/Statistical Learning/sclass.csv")


billboard <- read.csv("~/Desktop/UT/Statistical Learning/billboard.csv")

olympics_top20 <- read.csv("~/Desktop/UT/Statistical Learning/olympics_top20.csv")
```
<br>

### 1) Data visualization: flights at ABIA

<br>

**Does average airline departure delay vary by day of the week? Or maybe by time of the year?**


After averaging departure delays by day of week and month, we can understand how airline departure delays may vary. 
<br>
```{r ABIA setup, include=FALSE}
ABIA$UniqueCarrier <- gsub("EV", "ExpressJet", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("YV", "Mesa", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("B6", "Jetblue", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("OH", "Jetstream Intl", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("DL", "Delta", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("UA", "United", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("AA", "American", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("WN", "Southwest", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("CO", "NW Cargo Intl", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("OO", "Skywest", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("MQ", "Envoy Air", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("XE", "JSX Air", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("9E", "Endeavor Air", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("F9", "Frontier", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("US", "Puerto Rico Intl", ABIA$UniqueCarrier)
ABIA$UniqueCarrier <- gsub("NW", "Northwest", ABIA$UniqueCarrier)

by_carrier_daily = ABIA %>%
  group_by(UniqueCarrier, DayOfWeek)%>%
  summarize(mean_delay=mean(DepDelay, na.rm=TRUE))

by_carrier_monthly = ABIA %>%
  group_by(UniqueCarrier, Month)%>%
  summarize(mean_delay=mean(DepDelay, na.rm=TRUE))

delay_day_carrier=ggplot(by_carrier_daily) + 
  geom_col(aes(x=factor(DayOfWeek), y=mean_delay, color=DayOfWeek)) + 
  labs(x="Day of Week",
       y="Average Delay (minutes)",
       title= "Average Delays by Day of Week") +
  facet_wrap(facets= ~fct_reorder(UniqueCarrier, mean_delay, .desc = TRUE))

delay_month_carrier=ggplot(by_carrier_monthly) + 
  geom_col(aes(x=factor(Month), y=mean_delay, color=Month)) + 
  labs(x="Month",
       y="Average Delay (minutes)",
       title= "Average Delays by Month of Year")+
  facet_wrap(facets= ~fct_reorder(UniqueCarrier, mean_delay, .desc = TRUE))
```
<br>
```{r ABIA plot1, echo=FALSE, fig.cap="\\label{fig:figs}ExpressJet, Jetstream Intl, Mesa, and Jetblue are the four airline carriers with the longest average departure delays, while Endeavor Air, Northwest, Frontier, and Puerto Rico Intl have the shortest delays. Moreover, most carriers experience their longest delays at the beginning of the week or the weekend. Meanwhile, the day with shortest delays typically is Tuesday or Wednesday."}
plot(delay_day_carrier)
```

<br>
<br>
<br>
```{r ABIA plot2, echo=FALSE, fig.cap="\\label{fig:figs}Compared to the average daily delays, the same four airlines have the longest delays and the same four airlines have the shortest delays throughout the year. However, across most airlines we see the lowest delays around the fall. This makes sense when we consider weather and travel patterns throughout the year."}
plot(delay_month_carrier)
```


<br>

**From these results, we can conclude that traveling early-mid week in the fall on Endeavor Air, Frontier, Puerto Rico Intl, or Northwest airlines will most likely result in the shortest departure delays out of Austin.**
<br>
<br>

### 2) Wrangling the Billboard Top 100

<br>
The Billboard Top 100 helps us understand what artists and songs are popular throughout musical history. From Billboard Top 100 data, we can better understand top songs, musical diversity, and long-run hits.

<br>

#### Part A

<br>
```{r billboard, include=FALSE}
#part a 
top10 = billboard%>%
  group_by(performer, song)%>%
  summarize(count=n())%>%
  arrange(desc(count))%>%
  head(10)

#part b
chart_per_year = billboard%>%
  filter(year>=1959 & year<=2020)%>%
  group_by(year,song_id)%>%
  summarize(count=n())


musical_diversity_per_year = chart_per_year %>%
  group_by(year)%>%
  summarize(count=n())


musical_diversity_plot = ggplot(musical_diversity_per_year) +
  geom_line(aes(x=year, y=count)) +
  labs(x="Year",
       y="Number of Unique Songs",
       title="Musical Diversity of Billboard Top 100 by Year")

# part c 
ten_week_hit = billboard %>%
  filter(weeks_on_chart>=10)%>%
  group_by(performer, song_id)%>%
  summarize(count=n())


ten_week_hit_artist = ten_week_hit%>%
  group_by(performer)%>%
  summarize(count=n())%>% 
  filter(count>=30) %>%
  arrange(desc(count))


ten_week_hit_plot = ggplot(ten_week_hit_artist, aes(x=performer, y=count)) +
  geom_col(aes(x=performer, y=count)) +
  labs(x="Artist",
       y= "Number of Hits",
       title="Number of Ten-Week Hits per Artist") + geom_text(aes(label=count), hjust = -0.5) +
  coord_flip() 
```
#### Top 10 Songs on Billboard Top 100 by Total Number of Weeks
| **Performer**                             | **Song**                            | **Count** |
|-------------------------------------------|-------------------------------------|-----------|
| Imagine Dragons                           | Radioactive                         | 87        |
| AWOLNATION                                | Sail                                | 79        |
| Jason Mraz                                | Im Yours                            | 76        |
| The Weeknd                                | Blinding Lights                     | 76        |
| LeAnn Rimes                               | How Do I Live                       | 69        |
| LMFAO Featuring Lauren Bennett & GoonRock | Party Rock Anthem                   | 68        |
| OneRepublic                               | Counting Stars                      | 68        |
| Adele                                     | Rolling In The Deep                 | 65        |
| Jewel                                     | Foolish Games/You Were Meant For Me | 65        |
| Carrie Underwood                          | Before He Cheats                    | 64        |

<br>
The number one song is Radioactive by Imagine Dragons with 87 total weeks spent on the Billboard Top 100. Coming in at number ten is Before He Cheats by Carrie Underwood, with 64 weeks on the Billbaord Top 100.
<br>
<br>

#### Part B

<br>
```{r musical diversity, echo=FALSE, fig.cap="\\label{fig:figs}The musical diversity (or number of unique songs) of the Billboard Top 100 was highest in the 1960s and around 2020. From about 1980 to 2010, there was less musical diversity on the Billboard Top 100, which suggests that listeners preferred a less diverse collection of msuic during this time."}
plot(musical_diversity_plot)
```

<br>
<br>

#### Part C

<br>
```{r longhits, echo=FALSE, fig.cap="\\label{fig:figs}Only 19 aritsts had at least 30 songs on the Billboard Top 100 for at 10 weeks or longer. Of these 19 artists, we can see that the top three artists are: Elton John with 52 hits, Madonna with 44 hits, and Kenny Chesney with 42 hits."}
plot(ten_week_hit_plot)
```

<br>
<br>

## 3) Wrangling the Olympics

<br>

#### Part A

<br>

**The 95th percentile of heights for female competitors across all Athletics events**

<br>
```{r olympics parta, include=FALSE}
female_heights_95 = olympics_top20%>%
  filter(sex=="F" & sport=="Athletics")%>%
  group_by(event)%>%
  summarize(q95 = quantile(height, 0.95))
female_heights_95
```

| **Event**                                   | **95th Percentile** |
|---------------------------------------------|---------------------|
| Athletics Women's 1,500 metres              |              172.00 |
| Athletics Women's 10 kilometres Walk        |              170.00 |
| Athletics Women's 10,000 metres             |              167.80 |
| Athletics Women's 100 metres                |              179.60 |
| Athletics Women's 100 metres Hurdles        |              176.00 |
| Athletics Women's 20 kilometres Walk        |              173.00 |
| Athletics Women's 200 metres                |              180.00 |
| Athletics Women's 3,000 metres              |              170.00 |
| Athletics Women's 3,000 metres Steeplechase |              176.80 |
| Athletics Women's 4 x 100 metres Relay      |              176.00 |
| Athletics Women's 4 x 400 metres Relay      |              178.00 |
| Athletics Women's 400 metres                |              180.00 |
| Athletics Women's 400 metres Hurdles        |              181.40 |
| Athletics Women's 5,000 metres              |              170.75 |
| Athletics Women's 80 metres Hurdles         |              174.80 |
| Athletics Women's 800 metres                |              175.75 |
| Athletics Women's Discus Throw              |              188.40 |
| Athletics Women's Hammer Throw              |              183.90 |
| Athletics Women's Heptathlon                |              188.70 |
| Athletics Women's High Jump                 |              188.00 |
| Athletics Women's Javelin Throw             |              186.20 |
| Athletics Women's Long Jump                 |              181.35 |
| Athletics Women's Marathon                  |              171.40 |
| Athletics Women's Pentathlon                |              179.00 |
| Athletics Women's Pole Vault                |              183.00 |
| Athletics Women's Shot Put                  |              191.50 |
| Athletics Women's Triple Jump               |              186.90 |

<br>
<br>

#### Part B

<br>
**Which single women's event had the greatest variability in competitor's heights across the entire history of the Olympics, as measured by the standard deviation?**
<br>

```{r olympics partb, include=FALSE}
greatest_variation_height = olympics_top20%>%
  filter(sex=="F")%>%
  group_by(event)%>%
  summarize(sd= sd(height))%>%
  arrange(desc(sd))%>%
  head(1)
greatest_variation_height
```
| **Event**                  | **Standard Deviation** |
|----------------------------|------------------------|
| Rowing Women's Coxed Fours |               10.86549 |

<br>
The Women's Rowing Coxed Fours has the highest variablity in height across women's single Olympic events.
<br>
<br>

#### Part C

<br>
**How has the average age of Olympic swimmers changed over time?**
<br>
```{r, include=FALSE}
swimming_age = olympics_top20%>%
  filter(sport=="Swimming")%>%
  group_by(year)%>%
  summarize(mean_age = mean(age))
swimming_age

swimming_age_plot = ggplot(swimming_age)+
  geom_line(aes(x=year, y=mean_age)) + 
  labs(x="Year",
       y="Average Age",
       title="Average Age of Olympic Swimmers")
swimming_age_M_F = olympics_top20%>%
  filter(sport=="Swimming")%>%
  group_by(year, sex)%>%
  summarize(mean_age_M_F = mean(age))%>%
  arrange((mean_age_M_F))
swimming_age_M_F

swimming_age_M_F_df <- data.frame(swimming_age_M_F, stringsAsFactors = FALSE)
swimming_age_M_F_df

swimming_age_M_F_plot = ggplot(swimming_age_M_F_df) +
  geom_line(aes(x=year, y=mean_age_M_F, color=sex)) +
  labs(x="Year",
       y="Average Age",
       title="Average Age of Male and Female Olympic Swimmers")
swimming_age_M_F_plot
```
```{r swim agg plot, echo=FALSE, fig.cap="\\label{fig:figs}The average age of all Olympic swimmers in the early 1900s is around 27 years old. However, the average age dips to around 20 during the early and mid 1900s. By the 2010s, however, the average age is back up to around 23."}
plot(swimming_age_plot)
```

<br>
<br>
```{r swim age plot2, echo=FALSE, fig.cap="\\label{fig:figs}The average age for male Olympic swimmers starts around 16 years old in 1900, but shoots up to about 32 in 1925 before dropping back to about 20. By the 2010s, the average male age is about 24. Meanwhile, for female Olympic swimmers, the average age is starts around 16 in 1925 and stays on an upward trend to around 22 in the 2010s."}
plot(swimming_age_M_F_plot)
```

<br>
<br>

## 4) K-nearest Neighbors

```{r KNN, include=FALSE}
sclass_350 = sclass%>%
  filter(trim=="350")
sclass_350_split =  initial_split(sclass_350, prop=0.8)
sclass_350_train = training(sclass_350_split)
sclass_350_test  = testing(sclass_350_split)


#k_grid = c(2, 10, 15, 20, 40, 60, 80, 100)

knn2_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=2)
knn2_sclass350_rmse = rmse(knn2_sclass350, sclass_350_test)

knn10_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=10)
knn10_sclass350_rmse = rmse(knn10_sclass350, sclass_350_test)

knn15_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=15)
knn15_sclass350_rmse = rmse(knn15_sclass350, sclass_350_test)

knn20_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=20)
knn20_sclass350_rmse = rmse(knn20_sclass350, sclass_350_test)

knn40_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=40)
knn40_sclass350_rmse = rmse(knn40_sclass350, sclass_350_test)

knn60_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=60)
knn60_sclass350_rmse = rmse(knn60_sclass350, sclass_350_test)

knn80_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=80)
knn80_sclass350_rmse = rmse(knn80_sclass350, sclass_350_test)

knn100_sclass350 = knnreg(price ~ mileage, data=sclass_350_train, k=100)
knn100_sclass350_rmse = rmse(knn100_sclass350, sclass_350_test)

RMSE_K_sclass350 <- data.frame(K = c(2,10,15,20,40,60,80,100),
                         RMSE = c(knn2_sclass350_rmse, knn10_sclass350_rmse, knn15_sclass350_rmse, knn20_sclass350_rmse, knn40_sclass350_rmse, knn60_sclass350_rmse, knn80_sclass350_rmse, knn100_sclass350_rmse))


#plot RMSE versus K
RMSE_K_sclass350_plot = ggplot(RMSE_K_sclass350) +
  geom_point(aes(x=factor(K),y=RMSE)) +
  labs(title = "S Class 350: Out-of-sample RMSE for Values of K",
       x = "K",
       y = "Out-of-sample RMSE")
RMSE_K_sclass350_plot

#Fit the model to training set and make predictions on test set
sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn2 = predict(knn2_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn10 = predict(knn10_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn15 = predict(knn15_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn20 = predict(knn20_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn40 = predict(knn40_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn60 = predict(knn60_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn80 = predict(knn80_sclass350, sclass_350_test))

sclass_350_test =  sclass_350_test %>%
  mutate(price_pred_knn100 = predict(knn100_sclass350, sclass_350_test))

sclass_350_predictions = ggplot(data = sclass_350_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2) + 
  ylim(6000, 110000) +
  geom_line(aes(x = mileage, y = price_pred_knn40), color='red', size=1.5) +
  labs(x="Mileage",
       y="Price",
       title="S Class 350: Price & Mileage for K=40")
sclass_350_predictions 

############################
sclass_65 = sclass %>%
  filter(trim=="65 AMG")
sclass_65_split =  initial_split(sclass_65, prop=0.8)
sclass_65_train = training(sclass_65_split)
sclass_65_test  = testing(sclass_65_split)

#k_grid = c(2, 10, 15, 20, 40, 60, 80, 100)

knn2_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=2)
knn2_sclass65_rmse = rmse(knn2_sclass65, sclass_65_test)

knn10_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=10)
knn10_sclass65_rmse = rmse(knn10_sclass65, sclass_65_test)

knn15_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=15)
knn15_sclass65_rmse = rmse(knn15_sclass65, sclass_65_test)

knn20_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=20)
knn20_sclass65_rmse = rmse(knn20_sclass65, sclass_65_test)

knn40_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=40)
knn40_sclass65_rmse = rmse(knn40_sclass65, sclass_65_test)

knn60_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=60)
knn60_sclass65_rmse = rmse(knn60_sclass65, sclass_65_test)

knn80_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=80)
knn80_sclass65_rmse = rmse(knn80_sclass65, sclass_65_test)

knn100_sclass65 = knnreg(price ~ mileage, data=sclass_65_train, k=100)
knn100_sclass65_rmse = rmse(knn100_sclass65, sclass_65_test)

RMSE_K_sclass65 <- data.frame(K = c(2,10,15,20,40,60,80,100),
                               RMSE = c(knn2_sclass65_rmse, knn10_sclass65_rmse, knn15_sclass65_rmse, knn20_sclass65_rmse, knn40_sclass65_rmse, knn60_sclass65_rmse, knn80_sclass65_rmse, knn100_sclass65_rmse))


#plot RMSE versus K
RMSE_K_sclass65_plot = ggplot(RMSE_K_sclass65) +
  geom_point(aes(x=factor(K),y=RMSE)) +
  labs(title = "S Class 65 AMG: Out-of-sample RMSE for Values of K",
       x = "K",
       y = "Out-of-sample RMSE")
RMSE_K_sclass65_plot

#Fit the model to training set and make predictions on test set
sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn2 = predict(knn2_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn10 = predict(knn10_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn15 = predict(knn15_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn20 = predict(knn20_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn40 = predict(knn40_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn60 = predict(knn60_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn80 = predict(knn80_sclass65, sclass_65_test))

sclass_65_test =  sclass_65_test %>%
  mutate(price_pred_knn100 = predict(knn100_sclass65, sclass_65_test))

sclass_65_predictions = ggplot(data = sclass_65_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.2) + 
  ylim(6000, 110000) +
  geom_line(aes(x = mileage, y = price_pred_knn60), color='red', size=1.5) +
  labs(x="Mileage",
       y="Price",
       title="S Class 65 AMG: Price & Mileage for K=60")
sclass_65_predictions 
```
<br>

#### S CLass 350 

<br>

```{r sclass350, echo=FALSE, fig.cap="\\label{fig:figs}After plotting the RMSE for different values of K, it is clear the the RMSE bottoms out around K=40 for the S Class 350."}
plot(RMSE_K_sclass350_plot)
```

<br>
<br>
```{r sclass350_2, echo=FALSE,fig.cap="\\label{fig:figs}Using the optimal value of K=40 for the S Class 350, we can plot the fitted model of price against mileage."}
plot(sclass_350_predictions )
```

<br>
<br>

#### S Class 65 AMG 

<br>

```{r sclass65, echo=FALSE, fig.cap="\\label{fig:figs}After plotting the RMSE for different values of K, it is clear the the RMSE bottoms out around K=60 for the S Class 65 AMG."}
plot(RMSE_K_sclass65_plot)
```

<br>
<br>
```{r sclass65_2, echo=FALSE, warning=FALSE, fig.cap="\\label{fig:figs}Using the optimal value of K=60 for the S Class 65 AMG, we can plot the fitted model of price against mileage."}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
plot(sclass_65_predictions )
```

<br>
<br>
**Relative to the S Class 350, the S Class 65 AMG has a higher optimal value of K. The S Class 65 AMG may have a larger optimal value of K, because the predicted values are not as closely grouped together. Therefore, a larger K is necessary to get a more accurate estimate.**
<br>
<br>
<br>



